---
title: "Assignment 4"
subtitle: "PUBH 7462 - Assignment 4"
author: "Patrick Torralba"
format: 
  pdf:
    embed-resources: true
    prefer-html: true
execute: 
  echo: false
  message: false
  warning: false
editor: visual
---

## Introduction

```{r load atus_functions + data, echo=FALSE}
source("atus_functions.R")

atus <- read.csv("atus_raw-1.csv")

```

```{r library, echo=FALSE}
library(tidyverse)
library(showtext) # for ggplot
showtext_auto()
library(gtsummary) # for table
library(gt)
library(knitr)
library(kableExtra)
```

## Question

How does the amount of time spent sleeping per day vary by age group?

## Methods

I created a table to show the mean, median, and standard deviation of the number of hours spent on sleeping for different age groups.

And, I also created a simple bar graph that shows the average sleeping hours by age group.

## Results

```{r data cleaning, echo=FALSE}
sleep <- atus %>% 
  select(AGE, ACT_SLEEPING) %>% 
  mutate(age_group = case_when(AGE < 18 ~ "1-17",
                               AGE >= 18 & AGE < 25 ~ "18-24",
                               AGE >= 25 & AGE < 40 ~ "25-39",
                               AGE >= 39 & AGE < 60 ~ "40-59",
                               AGE >= 60 & AGE < 80 ~ "60-79",
                               AGE >= 80 & AGE < 100 ~ "80-100",
                               .default = "100+"),
         sleep_hours = round(ACT_SLEEPING / 60,1)) %>% 
  group_by(age_group) %>% 
  summarize(avg_sleep_hours = round(mean(sleep_hours, na.rm = TRUE),1),
            median_sleep_hours = round(median(sleep_hours, na.rm = TRUE),1),
            st.dev_sleep_hours = round(sd(sleep_hours, na.rm = TRUE),1))

```

```{r}
#| label: tbl-sleep 
#| tbl-cap: "Daily Sleep Summary"
#| echo: FALSE

sleep %>%
  kable(
    col.names = c("Age Group (Years)", "Average", "Median", "SD"),
    align = "lccc", 
    digits = 1) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  add_header_above(c(" " = 1, "Daily Sleep Summary (Hours)" = 3))
```

```{r}
#| label: fig-sleep 
#| fig-cap: "Average Sleep (Hours) by Age Group"
#| echo: FALSE

ggplot(sleep, aes(x = age_group, y = avg_sleep_hours, fill = age_group)) +
  geom_col() +
  geom_text(aes(label = round(avg_sleep_hours, 1)),
            hjust = -1.2,
            color = "black",
            fontface = "bold") +
  labs(x = NULL,
       y = "Average Sleep",
       fill = "Age Group",
       title = "Average Sleep (Hours) by Age Group") +
  coord_flip() +
  theme_minimal(base_family = "Verdana") +
  theme(axis.title.x = element_blank(),
        axis.ticks.length.x = unit(0, "cm"),
        axis.text.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = "none")

```

The group with the highest and lowest daily sleep hours are `r sleep$age_group[which.max(sleep$avg_sleep_hours)]` year olds and `r sleep$age_group[which.min(sleep$avg_sleep_hours)]` year olds, respectively. @fig-sleep

The median didn't vary too far from the mean, and the standard deviation for all groups is between 2.1-2.5 hours. @tbl-sleep

## Discussion

* 1-17 year age-group has the highest average daily sleep hours
* 40-59 year age-group has the lowest average daily sleep hours
* Average and median values are roughly similar

::: {.callout-important}

There are some limitations to this analysis. For example, babies and infants were grouped in the 1-17 age-group. This is a limitation because since babies and infants sleep more often than older children and adolescents, they can skew the average. Another limitation is that statistical significance was not tested. Therefore, it is unknown if the average sleeping hours by age-group is actually meaningful.

:::